Functional Analyses of Distinct Oxygen Concentrations and Light Niches of *Prochlorococcus marinus*

### 2024-02-14 Code/Manuscript_O2.Rmd
metaproteomics mat met

## Ocean Protein Portal analyses {.unnumbered}

@@

>>>>

## MetaProteomic data {.unnumbered}

The OceansMap Protein Portal (OPP; https://proteinportal.whoi.edu/) is an open access online data repository  (Woods Hole Oceanographic Institute, WHOI) for mass spectroscopy data on marine microbial peptides, found at various depths and locations worldwide. We screened the OPP for proteins annotated as from *Prochlorococcus* strains, to identify differential strategies employed by strains living at varying depths and oxygen levels within the marine water column.  We focussed particularly on the proteins from photosynthetic electron transport and photosystem repair proteins – Pcbs, PSI and II, Cytochrome b6f, NDH, PTOX, Rubsico, ATP Synthase and the FtsH proteases. the analyzed metaproteomic data was collected from 12 locations in the tropical North Pacific ocean along 150 W from 18 N of the equator between October 1, 2011 and October 25, 2011 during the voyage of the R/V Kilo Moana MetZyme cruise KM1128 (https://www.rvdata.us/search/cruise/KM1128; original datasets in the Biological and Chemical Oceanography Data Management Office repository; https://www.bco-dmo.org/project/2236).

### Sample Treatment and Peptide Identification
Collection and treatment of protein samples were performed following  Saito et al., (2014, 2015) XXXXADD to ZoteroXXXX . Briefly, samples of seawater from depths of 20, 40, 50, 60, 70, 80, 90, 120, 150, 200, 250, 300, 380, 400, 500, 550, 600, and 800 m below the ocean surface were pumped through a 0.2 µm filter, preserved in RNAlater and frozen at -80°C until extraction. Proteins were extracted from the filter in an SDS-based detergent, embedded in tube gel, alkylated and reduced prior to in-gel trypsin digestion. Peptide spectra were generated using a Q-Exactive Orbitrap Mass Spectrometer, searched in the SEQUEST CITATIONXXXX and labelled with the most likely protein and species annotation from Uniprot.  Oxygen levels at the location of sampling were recorded.

@@

### Data Processing and Bioinformatic Analyses

	Datasets were obtained from the KM1128 entry in the BCO-DMO database (https://www.bco-dmo.org/deployment/59053) accessed via the OPP in June 2019. Datasets contained: i) peptide sequences and sample identification (ID) number; ii) sample ID number, station, depth in meters below the surface the sample was collected at, best-hit BLASTP protein and species annotation and the corresponding Uniprot Entry number for the identified proteins and iii) depth, station and oxygen levels. Environmental conditions (depth and oxygen concentration) were joined to peptide sequence and BLASTP annotations by ID number, depth and station using the functions in the Tidyverse package (https://www.tidyverse.org/) in RStudio v1.2.5019 (https://rstudio.com/). The resulting merged dataset was filtered for Prochlorococcus peptides detected from 0 to 300 m below the surface that were annotated as a subunit of PSI, PSII, Pcbs, Cyt b6f, Rubisco, NDH, plastocyanin, ferredoxin, ATP Synthase or FtsH proteases. Peptides were re-annotated for consistency and labelled according to strain, clade, subunit and protein complex (if possible). Data for each strain was plotted against exact depth and O2 values.
All protein sequences were obtained from UniProt (https://www.uniprot.org/) and analyzed in Molecular Evolution and Genetic Analyses X (MEGAX) software  (https://www.megasoftware.net/). Sequences for each protein and/or subunits for each of the thirteen Prochlorococcus strains identified in the dataset were aligned with MUSCLE using UPGMA cluster method and a lambda of 24 with a -2.9 gap open penalty and 1.20 hydrophobicity multiplier.  Overall mean pairwise distance between protein sequences was determined using bootstrap variance estimation methods. Maximum likelihood phylogenetic trees were assembled using 1000 bootstrap replications with a 95% site coverage cut off. Prochlorococcus FtsH isoform identities were specified by phylogenetic comparison to the FtsH proteases of Synechocystis sp. PCC6803.

#### Data interpretation, assumptions and limitations
	Graphs depict the presence of mass-spectrally detected peptides at varying depths and O2 levels in the water column for each Prochlorococcus strain identified in the dataset. When analyzing the presence of a particular protein complex, all peptides belonging to all subunits were included to give the greatest number of data points. Where a particular metabolic pathway or reaction was examined, the activity of the pathway was interpreted by analysis of the rate-limiting enzyme or unique enzyme in the pathway (i.e. Rubisco was used as proxy for the activity of the Calvin Cycle). 
	Although not a direct measurement, it is assumed that light intensity is inversely correlated with depth. The terms “light” and “depth” therefore denote the same parameter the present paper. It is also assumed that the presence of a protein indicates that it is active under the given environmental parameters. 
	As this data was acquired on a discovery mission rather than through targeted peptide approaches, it is difficult to discern how accurate strain annotation of a particular peptide is, particularly as the proteins of interest in this study are highly conserved in sequence. We are confident in clade and ecotypic classification for each protein examined and although exact strain annotation may be ambiguous for the purposes of this research, we assume strain identification is accurate.
Another caveat to interpretation of this data is the “peptide detection bias” inherent to mass spectrometry: certain peptides are more easily detected and thus dominate the spectra over less easily detected peptides. The data is also limited by the number and nature of protein spectra in the SEQUEST database: a peptide sequence cannot be determined unless there is already a known spectrum for that peptide in the SEQUEST database, hence some peptides of interest may not be identifiable. Therefore, the absence of a particular does not necessarily rule out its presence overall. This bias, however, does not render the laws of probability moot: there is still likely to be a correlation between peptide abundance and detection. Furthermore, a peptide must be detected above a certain threshold abundance in order to be considered an accurate ‘hit’. Therefore, the ‘absence’ of a particular protein structure is here interpreted as low in quantity, while its presence is interpreted as high in quantity, and this is reflected in the results and discussion of the present data.
Finally, we have not assessed the ativity of these peptides, only their presence or absence from a particular environment. Therefore, we cannot accurately gauge total activity of Prochlorococcus proteins.
